---
output: 
  md_document:
    variant: gfm
  html_document:
    keep_md: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  message = FALSE
)
library(mvtnorm)
```

# An R Package for Density Ratio Estimation

#### *Koji MAKIYAMA (@hoxo_m)*

<!-- badges: start -->
[![Travis-CI Build Status](https://travis-ci.org/hoxo-m/densratio.svg?branch=master)](https://travis-ci.org/hoxo-m/densratio)
[![CRAN Version](http://www.r-pkg.org/badges/version/densratio)](https://CRAN.R-project.org/package=densratio)
[![CRAN Downloads](http://cranlogs.r-pkg.org/badges/densratio)](http://cranlogs.r-pkg.org/badges/densratio/)
[![Coverage Status](https://coveralls.io/repos/github/hoxo-m/densratio/badge.svg?branch=master)](https://coveralls.io/github/hoxo-m/densratio?branch=master)
[![Say Thanks!](https://img.shields.io/badge/Say%20Thanks-!-1EAEDB.svg)](https://saythanks.io/to/hoxo-m)
<!-- badges: end -->

## 1. Overview

**Density ratio estimation** is described as follows: for given two data samples `x` and `y` from unknown distributions `p(x)` and `q(y)` respectively, estimate `w(x) = p(x) / q(x)`, where `x` and `y` are d-dimensional real numbers.

The estimated density ratio function `w(x)` can be used in many applications such as **anomaly detection** [Hido et al. 2011], **changepoint detection** [Liu et al. 2013], and **covariate shift adaptation** [Sugiyama et al. 2007].
Other useful applications about density ratio estimation were summarized by [Sugiyama et al. 2012].

The package **densratio** provides a function `densratio()`.
The function outputs an object that has a function to estimate density ratio.

For example, 

```{r result, cache=TRUE}
set.seed(3)
x <- rnorm(200, mean = 1, sd = 1/8)
y <- rnorm(200, mean = 1, sd = 1/2)

library(densratio)
result <- densratio(x, y)
```

The function `densratio()` estimates the density ratio of `p(x)` to `q(y)`, `w(x) = p(x)/q(y) = Norm(1, 1/8) / Norm(1, 1/2)`, and provides a function to compute estimated density ratio. 
The result object has a function `compute_density_ratio()` that can compute the estimated density ratio `w_hat(x)` for any d-dimensional input `x` (now d=1).

```{r compute-estimated-density-ratio, fig.width=5, fig.height=4}
new_x <- seq(0, 2, by = 0.06)
w_hat <- result$compute_density_ratio(new_x)

plot(new_x, w_hat, pch=19)
```

In this case, the true density ratio `w(x) = p(x)/q(y) = Norm(1, 1/8) / Norm(1, 1/2)` can be computed precisely. 
So we can compare `w(x)` with the estimated density ratio `w_hat(x)`.

```{r compare-true-estimate, fig.width=5, fig.height=4}
true_density_ratio <- function(x) dnorm(x, 1, 1/8) / dnorm(x, 1, 1/2)

plot(true_density_ratio, xlim=c(-1, 3), lwd=2, col="red", xlab = "x", ylab = "Density Ratio")
plot(result$compute_density_ratio, xlim=c(-1, 3), lwd=2, col="green", add=TRUE)
legend("topright", legend=c(expression(w(x)), expression(hat(w)(x))), col=2:3, lty=1, lwd=2, pch=NA)
```

## 2. Installation

You can install the **densratio** package from [CRAN](https://CRAN.R-project.org/package=densratio).

```{r eval=FALSE}
install.packages("densratio")
```

You can also install the package from [GitHub](https://github.com/hoxo-m/densratio).

```{r eval=FALSE}
install.packages("devtools") # if you have not installed "devtools" package
devtools::install_github("hoxo-m/densratio")
```

The source code for **densratio** package is available on GitHub at

- https://github.com/hoxo-m/densratio.

## 3. Details

`densratio()` has `method` argument that you can pass `"uLSIF"`, `"RuSLIF"`, or `"KLIEP"`.

- **uLSIF** (unconstrained Least-Squares Importance Fitting) is the default method.
This algorithm estimates density ratio by minimizing the squared loss.
You can find more information in [Kanamori et al. 2009] and [Hido et al. 2011].

- **RuLSIF** (Relative unconstrained Least-Squares Importance Fitting).
This algorithm estimates relative density ratio by minimizing the squared loss.
You can find more information in [Yamada et al. 2011] and [Liu et al. 2013].

- **KLIEP** (Kullback-Leibler Importance Estimation Procedure).
This algorithm estimates density ratio by minimizing Kullback-Leibler divergence.
You can find more information in [Sugiyama et al. 2007].

There is a vignette for the package. For more detail, read it.

```{r open-vignette, eval=FALSE}
vignette("densratio")
```

You can also find it on CRAN.

- [An R Package for Density Ratio Estimation](https://CRAN.R-project.org/package=densratio/vignettes/densratio.html)

## 4. Related work

We have also developed a Python package for density ratio estimation.

- https://github.com/hoxo-m/densratio_py

The package is available on PyPI (Python Package Index).

- https://pypi.python.org/pypi/densratio

## 5. References

- Hido, S., Y. Tsuboi, H. Kashima, M. Sugiyama, and T. Kanamori.
**Statistical outlier detection using direct density ratio estimation.**
Knowledge and Information Systems, 2011.

- Kanamori, T., S. Hido, and M. Sugiyama.
**A Least-squares Approach to Direct Importance Estimation.**
Journal of Machine Learning Research, 2009.

- Liu, S., M. Yamada, N. Collier, M. Sugiyama.
**Change-point detection in time-series data by relative density-ratio estimation.**
Neural Net, 2013

- Sugiyama, M., S. Nakajima, H. Kashima, P. von Bünau, and M. Kawanabe. 
**Direct importance estimation with model selection and its application to covariate shift adaptation.**
NIPS 2007.

- Sugiyama, M., T. Suzuki, and T. Kanamori. 
**Density Ratio Estimation in Machine Learning.**
Cambridge University Press, 2012.

- Yamada, M., T. Suzuki, T. Kanamori, H. Hachiya, and M. Sugiyama.
**Relative Density-Ratio Estimation for Robust Distribution Comparison.**
NIPS 2011.
