---
title: "An R Package for Density Ratio Estimation"
author: Koji MAKIYAMA (@hoxo_m)
output: 
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(mvtnorm)
```

[![Travis-CI Build Status](https://travis-ci.org/hoxo-m/densratio.svg?branch=master)](https://travis-ci.org/hoxo-m/densratio)
[![CRAN Version](http://www.r-pkg.org/badges/version/densratio)](http://cran.rstudio.com/web/packages/densratio)
[![CRAN Downloads](http://cranlogs.r-pkg.org/badges/densratio)](http://cranlogs.r-pkg.org/badges/densratio)

## 1. Overview

**Density ratio estimation** is described as follows: for given two data samples `x` and `y` from unknown distributions `p(x)` and `q(y)` respectively, estimate `w(x) = p(x) / q(x)`, where `x` and `y` are d-dimensional real numbers.

The estimated density ratio function `w(x)` can be used in many applications such as the inlier-based outlier detection [1] and covariate shift adaptation [2].
Other useful applications about density ratio estimation were summarized by Sugiyama et al. (2012) [3].

The package **densratio** provides a function `densratio()` that returns a result has the function to estimate density ratio `compute_density_ratio()`.

For example, 

```{r result, cache=TRUE}
set.seed(3)
x <- rnorm(200, mean = 1, sd = 1/8)
y <- rnorm(200, mean = 1, sd = 1/2)

library(densratio)
result <- densratio(x, y)
result
```

In this case, the true density ratio `w(x)` is known, so we can compare `w(x)` with the estimated density ratio `w-hat(x)`.

```{r}
true_density_ratio <- function(x) dnorm(x, 1, 1/8) / dnorm(x, 1, 1/2)
estimated_density_ratio <- result$compute_density_ratio

plot(true_density_ratio, xlim=c(-1, 3), lwd=2, col="red", xlab = "x", ylab = "Density Ratio")
plot(estimated_density_ratio, xlim=c(-1, 3), lwd=2, col="green", add=TRUE)
legend("topright", legend=c(expression(w(x)), expression(hat(w)(x))), col=2:3, lty=1, lwd=2, pch=NA)
```

## 2. How to Install

You can install the **densratio** package from [CRAN](https://cran.r-project.org/web/packages/densratio/index.html).

```{r eval=FALSE}
install.packages("densratio")
```

You can also install the package from [GitHub](https://github.com/hoxo-m/densratio).

```{r eval=FALSE}
install.packages("devtools") # if you have not installed "devtools" package
devtools::install_github("hoxo-m/densratio")
```

The source code for **densratio** package is available on GitHub at

- https://github.com/hoxo-m/densratio.

## 3. Details

### 3.1. Basics

The package provides `densratio()` that the result has the function to estimate density ratio.

For data samples `x` and `y`,

```{r eval=FALSE}
library(densratio)

x <- rnorm(200, mean = 1, sd = 1/8)
y <- rnorm(200, mean = 1, sd = 1/2)

result <- densratio(x, y)
```

In this case, `result$compute_density_ratio()` can compute estimated density ratio.  

```{r}
w_hat <- result$compute_density_ratio(y)
plot(y, w_hat)
```

### 3.2. Methods

`densratio()` has `method` parameter that you can pass `"uLSIF"` or `"KLIEP"`.

- **uLSIF** (unconstrained Least-Squares Importance Fitting) is the default method.
This algorithm estimates density ratio by minimizing the squared loss.
You can find more information in Hido et al. (2011) [1].

- **KLIEP** (Kullback-Leibler Importance Estimation Procedure) is the anothor method.
This algorithm estimates density ratio by minimizing Kullback-Leibler divergence.
You can find more information in Sugiyama et al. (2007) [2].

The both methods assume that the denity ratio is represented by linear model: 

`w(x) = alpha1 * K(x, c1) + alpha2 * K(x, c2) + ... + alphab * K(x, cb)`

where `K(x, c) = exp(- ||x - c||^2 / (2 * sigma ^ 2))` is the Gaussian RBF.

`densratio()` performs the two main jobs:

- First, deciding kernel parameter `sigma` by cross validation,
- Second, optimizing kernel weights `alpha`.

As the result, you can obtain `compute_density_ratio()`.

### 3.3. Result and Paremeter Settings

`densratio()` outputs the result like as follows:

```{r echo=FALSE}
result
```

- **Kernel type** is fixed by Gaussian RBF.
- The **number of kernels** is the number of kernels in the linear model. You can change by setting `kernel_num` parameter. In default, `kernel_num = 100`.
- **Bandwidth(sigma)** is the Gaussian kernel bandwidth. In default, `sigma = "auto"`, the algorithms automatically select the optimal value by cross validation. If you set `sigma` a number, that will be used. If you set a numeric vector, the algorithms select the optimal value in them by cross validation.
- **Centers** are centers of Gaussian kernels in the linear model. These are selected at random from the data sample `x` underlying a numerator distribution `p(x)`. You can find the whole values in `result$kernel_info$centers`. 
- **Kernel weights** are alpha parameters in the linear model. It is optimaized by the algorithms. You can find the whole values in `result$alpha`. 
- **The funtion to estimate density ratio** is named `compute_density_ratio()`.

## 4. Multi Dimensional Data Samples

In the above, the input data samples `x` and `y` were one dimensional.
`densratio()` allows to input multidimensional data samples as `matrix`.

For example,

```{r cache=TRUE}
library(densratio)
library(mvtnorm)

set.seed(71)
x <- rmvnorm(300, mean = c(1, 1), sigma = diag(1/8, 2))
y <- rmvnorm(300, mean = c(1, 1), sigma = diag(1/2, 2))

result <- densratio(x, y)
result
```

Also in this case, we can compare the true density ratio with the estimated density ratio.

```{r fig.height=4}
true_density_ratio <- function(x) {
  dmvnorm(x, mean = c(1, 1), sigma = diag(1/8, 2)) /
    dmvnorm(x, mean = c(1, 1), sigma = diag(1/2, 2))
}
estimated_density_ratio <- result$compute_density_ratio

N <- 20
range <- seq(0, 2, length.out = N)
input <- expand.grid(range, range)
z_true <- matrix(true_density_ratio(input), nrow = N)
z_hat <- matrix(estimated_density_ratio(input), nrow = N)

par(mfrow = c(1, 2))
contour(range, range, z_true, main = "True Density Ratio")
contour(range, range, z_hat, main = "Estimated Density Ratio")
```

The dimensions of `x` and `y` must be same.

## 5. References

[1] Hido, S., Tsuboi, Y., Kashima, H., Sugiyama, M., & Kanamori, T. 
**Statistical outlier detection using direct density ratio estimation.**
Knowledge and Information Systems 2011. 

[2] Sugiyama, M., Nakajima, S., Kashima, H., von Bünau, P. & Kawanabe, M. 
**Direct importance estimation with model selection and its application to covariate shift adaptation.** NIPS 2007.

[3] Sugiyama, M., Suzuki, T. & Kanamori, T. 
**Density Ratio Estimation in Machine Learning.**
Cambridge University Press 2012.
