---
title: "An R Package for Density Ratio Estimation"
author: "Koji MAKIYAMA (@hoxo_m)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{An R Package for Density Ratio Estimation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(mvtnorm)
```

## 1. Overview

**Density ratio estimation** is described as follows: for given two data samples $x$ and $y$ from unknown distributions $p(x)$ and $q(y)$ respectively, estimate

$$
w(x) = \frac{p(x)}{q(x)}
$$

where $x$ and $y$ are $d$-dimensional real numbers.

The estimated density ratio function $w(x)$ can be used in many applications such as anomaly detection [1] and covariate shift adaptation [2].
Other useful applications about density ratio estimation were summarized by Sugiyama et al. (2012) [3].

The package **densratio** provides a function `densratio()`.
The function outputs an object that has a function to estimate density ratio.

For example, 

```{r}
set.seed(3)
x <- rnorm(200, mean = 1, sd = 1/8)
y <- rnorm(200, mean = 1, sd = 1/2)

library(densratio)
result <- densratio(x, y)
```

The function `densratio()` estimates the density ratio of $p(x)$ to $q(y)$,
$$
w(x) = \frac{p(x)}{q(y)} = \frac{\rm{Norm}(1, 1/8)}{\rm{Norm}(1, 1/2)}
$$
and provides a function to compute estimated density ratio.
The `result` object has a function `compute_density_ratio()` that can compute the estimated density ratio $\hat{w}(x) \simeq p(x)/q(y)$ for any $d$-dimensional input $x$ (now $d=1$).

```{r fig.width=5, fig.height=4}
new_x <- seq(0, 2, by = 0.05)
w_hat <- result$compute_density_ratio(new_x)

plot(new_x, w_hat, pch=19)
```

In this case, the true density ratio $w(x) = p(x)/q(y) = \rm{Norm}(1, 1/8) / \rm{Norm}(1, 1/2)$ can be computed precisely.
So we can compare $w(x)$ with the estimated density ratio $\hat{w}(x)$.

```{r fig.width=5, fig.height=4}
true_density_ratio <- function(x) dnorm(x, 1, 1/8) / dnorm(x, 1, 1/2)

plot(true_density_ratio, xlim=c(-1, 3), lwd=2, col="red", xlab = "x", ylab = "Density Ratio")
plot(result$compute_density_ratio, xlim=c(-1, 3), lwd=2, col="green", add=TRUE)
legend("topright", legend=c(expression(w(x)), expression(hat(w)(x))), col=2:3, lty=1, lwd=2, pch=NA)
```

## 2. How to Install

You can install the **densratio** package from [CRAN](https://cran.r-project.org/web/packages/densratio/index.html).

```{r eval=FALSE}
install.packages("densratio")
```

You can also install the package from [GitHub](https://github.com/hoxo-m/densratio).

```{r eval=FALSE}
install.packages("devtools") # If you have not installed "devtools" package
devtools::install_github("hoxo-m/densratio")
```

The source code for **densratio** package is available on GitHub at

- https://github.com/hoxo-m/densratio.

## 3. Details

### 3.1. Basics

The package **densratio** provides a function `densratio()`.
The function outputs an object that has a function to estimate density ratio.

For data samples `x` and `y`,

```{r eval=FALSE}
library(densratio)

x <- rnorm(200, mean = 1, sd = 1/8)
y <- rnorm(200, mean = 1, sd = 1/2)

result <- densratio(x, y)
```

Here, `result$compute_density_ratio()` is the function to compute estimated density ratio.

```{r fig.width=5, fig.height=4}
new_x <- seq(0, 2, by = 0.05)
w_hat <- result$compute_density_ratio(new_x)

plot(new_x, w_hat, pch=19)
```

### 3.2. Methods

`densratio()` has `method` argument that you can pass `"uLSIF"` or `"KLIEP"`.

- **uLSIF** (unconstrained Least-Squares Importance Fitting) is the default method.
This algorithm estimates density ratio by minimizing the squared loss.
You can find more information in Hido et al. (2011) [1].

- **KLIEP** (Kullback-Leibler Importance Estimation Procedure) is the anothor method.
This algorithm estimates density ratio by minimizing Kullback-Leibler divergence.
You can find more information in Sugiyama et al. (2007) [2].

The both methods assume that density ratio are represented by linear model

$$
w(x) = \alpha_1 K(x, c_1) + \alpha_2 K(x, c_2) + ... + \alpha_b K(x, c_b)
$$

where

$$
K(x, c) = \exp\left(-\frac{\|x - c\|^2}{2 \sigma ^ 2}\right)
$$

is the Gaussian RBF.

`densratio()` performs two main jobs:

- First, decide kernel parameter $\sigma$ by cross validation.
- Second, find the optimal kernel weights $\alpha_i$ (in other words, find the optimal coefficients of the linear model).

$\sigma$ and $\alpha_i$ are saved into result objects of `densratio()`, and used to compute estimated density ratio in `compute_density_ratio()`.

### 3.3. Result and Arguments

You can `print()` result objects of `densratio()` to see information.
Moreover, you can change some conditions to specify arguments of `densratio()`.

```{r}
print(result)
```

- **Kernel type** is fixed by Gaussian RBF.
- **Number of kernels** is the number of kernels in the linear model. You can change by setting `kernel_num` argument. In default, `kernel_num = 100`.
- **Bandwidth(sigma)** is the Gaussian kernel bandwidth. In default, `sigma = "auto"`, the algorithms automatically select the optimal value by cross validation. If you set `sigma` a single number, it will be used. If you set a numeric vector, the algorithms select the optimal value in them by cross validation.
- **Centers** are centers of Gaussian kernels in the linear model. These are selected at random from the data sample `x` underlying a numerator distribution $p(x)$. You can find the whole values in `result$kernel_info$centers`. 
- **Kernel weights** are the alpha parameters in the linear model. They have been optimaized by the algorithms. You can find the whole values in `result$alpha`. 
- **The funtion to estimate density ratio** is named `compute_density_ratio()`.

## 4. Multi Dimensional Data Samples

So far, the input data samples `x` and `y` were one dimensional.
`densratio()` allows to input multidimensional data samples as `matrix`.

For example,

```{r}
library(densratio)
library(mvtnorm)

set.seed(71)
x <- rmvnorm(300, mean = c(1, 1), sigma = diag(1/8, 2))
y <- rmvnorm(300, mean = c(1, 1), sigma = diag(1/2, 2))

result <- densratio(x, y)
result
```

Also in this case, we can compare the true density ratio with the estimated density ratio.

```{r fig.width=7, fig.height=4}
true_density_ratio <- function(x) {
  dmvnorm(x, mean = c(1, 1), sigma = diag(1/8, 2)) /
    dmvnorm(x, mean = c(1, 1), sigma = diag(1/2, 2))
}

N <- 20
range <- seq(0, 2, length.out = N)
input <- expand.grid(range, range)
w_true <- matrix(true_density_ratio(input), nrow = N)
w_hat <- matrix(result$compute_density_ratio(input), nrow = N)

par(mfrow = c(1, 2))
contour(range, range, w_true, main = "True Density Ratio")
contour(range, range, w_hat, main = "Estimated Density Ratio")
```

The dimensions of `x` and `y` must be same.

## 5. References

[1] Hido, S., Tsuboi, Y., Kashima, H., Sugiyama, M., & Kanamori, T. 
**Statistical outlier detection using direct density ratio estimation.**
Knowledge and Information Systems 2011. 

[2] Sugiyama, M., Nakajima, S., Kashima, H., von Bünau, P. & Kawanabe, M. 
**Direct importance estimation with model selection and its application to covariate shift adaptation.** NIPS 2007.

[3] Sugiyama, M., Suzuki, T. & Kanamori, T. 
**Density Ratio Estimation in Machine Learning.**
Cambridge University Press 2012.
